---
title: "Figure for discriminant analysis"
author: "Maxime Delmas"
date: "2023-02-22"
output: html_document
---
 
# libraries
```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(effsize)
library(ggpubr)
library(grid)
```

# Utility functions
```{r}
# Get divergence data in longer format
get_longer_data <- function(path){
  
  data_JS_divergences <- read_tsv(path)
  
  longer_data_JS_divergences <- data_JS_divergences %>% pivot_longer(cols = c("PosxPos", "NegxNeg", "PosxNeg"), names_to = "Type", values_to = "JSD")
  
  # remove lines with na in cases of compution errors (i can happen on some lines ...)
  longer_data_JS_divergences <- longer_data_JS_divergences[! is.na(longer_data_JS_divergences$JSD), ]
}
```


## Paths

```{r}
data_directory <- "../../output/np/discriminant-analysis/" # to change depending on file system
```


```{r}
t1.path <- paste(data_directory, "JS_divergence_eval_1_5000.tsv", sep="/")  # to change depending on file system

t1 <- get_longer_data(t1.path)
t1$Task <- "task1"

t2.path <- paste(data_directory, "JS_divergence_eval_2_5000.tsv", sep="/") # to change depending on file system

t2 <- get_longer_data(t2.path)
t2$Task <- "task2"

t <- rbind(t1, t2)

t$Template <- t$Template + 1
```


## Compute Mann Whitney Wilcoxon Test (also called the Wilcoxon Rank Sum Test)
```{r}

# prepare data
Wilcoxon_data_ <- t %>% select(model, Template, Task) %>% distinct()

Wilcoxon_data <- rbind((Wilcoxon_data_ %>% mutate(Compare = "VS-PosxPos")), (Wilcoxon_data_ %>% mutate(Compare = "VS-NegxNeg")))

Wilcoxon_data$CliffDelta <- 0
Wilcoxon_data$p.value <- 0

for (i in 1:nrow(Wilcoxon_data)) {
  model <- Wilcoxon_data[i, ]$model
  template <- Wilcoxon_data[i, ]$Template
  comparison <- Wilcoxon_data[i, ]$Compare
  data_wilcoxon <- t[t$model == model & t$Template == template, ]
  if(comparison == "VS-PosxPos"){
    
    wilcoxon_rank_sum_test <- wilcox.test(data_wilcoxon[data_wilcoxon$Type == "PosxNeg", ]$JSD, data_wilcoxon[data_wilcoxon$Type == "PosxPos", ]$JSD, paired = F, alternative = "greater")
    CliffD <- cliff.delta(data_wilcoxon[data_wilcoxon$Type == "PosxNeg", ]$JSD, data_wilcoxon[data_wilcoxon$Type == "PosxPos", ]$JSD)
  }
  else{
    
    wilcoxon_rank_sum_test <- wilcox.test(data_wilcoxon[data_wilcoxon$Type == "PosxNeg", ]$JSD, data_wilcoxon[data_wilcoxon$Type == "NegxNeg", ]$JSD, paired = F, alternative = "greater")
    CliffD <- cliff.delta(data_wilcoxon[data_wilcoxon$Type == "PosxNeg", ]$JSD, data_wilcoxon[data_wilcoxon$Type == "NegxNeg", ]$JSD)
    
  }
  Wilcoxon_data[i, ]$CliffDelta <- as.numeric(CliffD$estimate)
  Wilcoxon_data[i, ]$p.value <- as.numeric(wilcoxon_rank_sum_test$p.value)

}

DT::datatable(Wilcoxon_data)
```

What is the cliff-delta: It is the difference between the probability that X > Y and Y > X. In our case, it can be interpreted as the difference in the probabilities of observing a higher Jensen-Shanon divergence when comparing a positive statement against a negative statement, rather than when comparing two positive or negative statements.

When comparing 2 variables $X$ and $Y$:

$Delta = \frac{1}{mn} \sum_{i=1}^m \sum_{j=1}^n \delta_{ij}$ avec

$$
\delta_{ij} = 
  \begin{cases}
  1 \text{ si } x_i > y_j \\
  -1 \text{ si } x_i < y_j \\
  0 sinon
  \end{cases}
$$ 


Guidelines for interpreting the Cliff delta: https://sci-hub.st/https://doi.org/10.3102/10769986025002101
(Table 1)

```{r}
print(paste("Pourcentage de Cliff-delta < 0.11: ", sum(Wilcoxon_data$CliffDelta < 0.11) / nrow(Wilcoxon_data) * 100))
```

Figure:

```{r}

# New facet label names 
Wilcoxon_data$Compare <- factor(Wilcoxon_data$Compare, levels = c("VS-PosxPos", "VS-NegxNeg"), labels = c("D(P,N) > D(P,P)", "D(P,N) > D(N,N)"))
Wilcoxon_data$Task <- factor(Wilcoxon_data$Task, levels = c("task1", "task2"), labels = c("Fungus produces Chemical", "Chemical has antimicrobial activity"))

plot1 <- ggplot(Wilcoxon_data[Wilcoxon_data$Compare == "D(P,N) > D(P,P)", ], aes(x = factor(Template), y = model, fill = CliffDelta)) + 
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  # geom_text(aes(label = round(vs_Neg, 2)), color = "white", size = 6) +
  scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000",
                       limits = c(-1,1)) +
  facet_grid(. ~ Task, scales = "free_x", space = "free_x") +
  # ylab("Pre-trained Language Models") +
  theme(axis.text.y = element_text(size = 10),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


plot2 <- ggplot(Wilcoxon_data[Wilcoxon_data$Compare == "D(P,N) > D(N,N)", ], aes(x = factor(Template), y = model, fill = CliffDelta)) + 
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  # geom_text(aes(label = round(vs_Neg, 2)), color = "white", size = 6) +
  scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000",
                       limits = c(-1,1)) +
  facet_grid(. ~ Task, scales = "free_x", space = "free_x") +
  # ylab("Pre-trained Language Models") +
  theme(axis.text.y = element_text(size = 10),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


```{r}
plot(plot1)
```


```{r}
plot(plot2)
```

Pour la corrélation top 1

```{r}

t1.cor.path <- paste(data_directory, "top_1_cor_eval_1.tsv")
t1_top1_cor <- read_tsv(t1.cor.path)
t1_top1_cor$Task <- "task1"


t2.cor.path <- paste(data_directory, "top_1_cor_eval_2.tsv")
t2_top1_cor <- read_tsv(t2.cor.path)
t2_top1_cor$Task <- "task2"

top1_cor <- rbind(t1_top1_cor, t2_top1_cor)
top1_cor$Template <- top1_cor$Template + 1


top1_cor$Task <- factor(top1_cor$Task, levels = c("task1", "task2"), labels = c("Fungus produces Chemical", "Chemical has antimicrobial activity"))

```

```{r}

plot3 <- ggplot(top1_cor, aes(x = factor(Template), y = model, fill = cors)) + 
  geom_tile(color = "white",
            lwd = 1.5,
            linetype = 1) +
  # geom_text(aes(label = round(cors, 2)), color = "white", size = 6) +
  scale_fill_gradient2(low = "#075AFF",
                       high = "#FF0000",
                       limits = c(0,1)) +
  facet_grid(. ~ Task, scales = "free_x", space = "free_x") +
  # ylab("Pre-trained Language Models") +
  # xlab("Template index") +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x = element_text(size = 10), 
        axis.text.y = element_text(size = 10),
        strip.background = element_blank(),
        strip.text.x = element_blank())

```


```{r}
plot(plot3)
```


# Merge the 2 plots:

```{r}
full <- ggarrange(plot1, plot2, plot3, 
                  labels = c("A", "B", "C"),
                  font.label = list(size = 25, color = "black", face = "bold", family = NULL), 
                  common.legend = FALSE, 
                  legend = "right", ncol = 1)
```

```{r}
plot(full)
```