---
title: "Figures and tables for prediction analysiss"
author: "Maxime Delmas"
date: "2023-02-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(PTXQC)
library(rjson)
library(effsize)
library(ggpubr)
library(grid)
```

## functions

```{r}

read_predictions <- function(path){
  json.str <- readLines(path, ok = T, warn = F)
  json.data <- fromJSON(json.str[1])
  return(json.data)
}

# From the json data, creates the data.frame, with columns: subject.id, preds.label, preds.probas, rank
extract_prediction_proba_pairs <- function(json.data){
  
  # Get length
  N <- length(json.data[["result"]])
  M <- length(json.data[["result"]][[1]][["preds"]])
  
  # inti the dataset
  dataset <- data.frame(subject.id = vector("character", N * M), preds.label = vector("character", N * M), preds.probas = vector("numeric", N * M), preds.rank = rep(c(1:M), N))
  
  for(i in 1:N) {
    
    # get the data and fill to the corresponding indexes.
    result <- json.data[["result"]][[i]]
    
    splited.uuid <- strsplit(result[["uuid"]], "-")[[1]]
    subject.id <- splited.uuid[1]
    
    # preds
    preds <- unlist(result[["preds"]])
    preds.label <- preds[seq(1, length(preds), 2)]
    preds.probas <- preds[seq(2, length(preds), 2)]
    
    is.correct <- rep(FALSE, M)
    if(length(result[["corrected_preds"]]) > 0) {
      # correct preds
      correct.preds <- unlist(result[["corrected_preds"]])
      correct.preds.label <- correct.preds[seq(1, length(correct.preds), 2)]
      correct.preds.rank <- as.numeric(correct.preds[seq(2, length(correct.preds), 2)])
      
      # test if preds in correct preds
      is.correct <- preds.label %in% correct.preds.label
    }
    
    dataset[(M*(i - 1) + 1):(M * i ), "subject.id"] <- subject.id
    dataset[(M*(i - 1) + 1):(M * i ), "preds.label"] <- preds.label
    dataset[(M*(i - 1) + 1):(M * i ), "preds.probas"] <- as.numeric(preds.probas)
    dataset[(M*(i - 1) + 1):(M * i ), "is.correct"] <- is.correct
    
  }
  return(dataset)
}



create_plot <- function(big.df, ref){

      sub.df <- big.df[big.df$ref == ref, ]
    data_for_plot <- sub.df %>% mutate(masked.rank.bin = cut(rank.in.masked, breaks = c(-1, 1, 2, 5, 10, 20, 50), labels = c('exclusive', "top1", "top [2-5[", "top [5-10[", "[10-20[", "[20-50["), include.lowest = T, right = F))
    data_for_plot <- data_for_plot %>% group_by(ref, preds.rank, masked.rank.bin) %>% summarise(n = n())
    data_for_plot <- data_for_plot %>% group_by(ref, preds.rank) %>% mutate(prop = (n/sum(n)) * 100)
    data_for_plot$masked.rank.bin <- factor(data_for_plot$masked.rank.bin, levels = c('exclusive', "top1", "top [2-5[", "top [5-10[", "[10-20[", "[20-50["))
    
    p <- data_for_plot %>% ggplot(aes(x = preds.rank, y = prop, fill = masked.rank.bin)) +
      geom_bar(stat="identity", width=1) +
      theme_classic() +
      scale_fill_discrete(drop=FALSE) +
      theme(axis.text = element_text(size = 20, , colour="black"), text = element_text(size = 15, colour="black")) +
      xlab("") + 
      ylab("")
      # ggtitle(paste("property= ", ref)) +
    return(p)

}

get_objectlabels_freq_in_dataset <- function(path){

  lines <- readLines(path, ok = T, warn = F)
  
  N <- length(lines)
  print(paste("Number of examples:", N))
  v_obj.labels <- c()
  
  for(index in 1:N){
    
    line <- lines[index]
    
    json.data <- fromJSON(line)
    
    v_obj.labels <- c(v_obj.labels, json.data[["obj_labels"]])
    
  }
  out <- data.frame(obj.label = v_obj.labels)
  out <- out %>% group_by(obj.label) %>% summarise(n = n(), f = (n()/N) * 100 )
  return(out)
}


compute_pred_LCP <- function(x, y) {
    # Remove punctation or digit characters (if chemical for instance):
    process_str <- function(x) {
      tolower(gsub('[[:punct:][:digit:]]+','',x))
    }
    x.processed <- process_str(x)
    y.processed <- sapply(y, FUN = process_str, simplify = T, USE.NAMES = F)
    
    longest <- 0
    # cut x by space
    for (ind.x in strsplit(x.processed, " ")[[1]]) {
      # print(ind.x)
      # also test on all y cuted by space
      for(ind.y in strsplit(y.processed, " ")[[1]]){
        # test if longest prefix is
        l <- nchar(longestCommonPrefix(c(ind.x, ind.y)))
        if(l > longest){
          longest <- l
        }
      }
    }
    return(longest)
  }

get_table_labels <- function(path){
  
  lines <- readLines(path, ok = T, warn = F)
  
  N <- length(lines)
  
  out <- data.frame("subject.id" = vector("character", N), "subject.label" = vector("character", N))

  for(index in 1:N){
    
    line <- lines[index]
    
    json.data <- fromJSON(line)
    
    subject.id <- json.data[["sub_uri"]]
    
    # On ne considère pas les synonymes car ils ne sont pas envoyé au modèle.
    subject.label <- json.data[["sub_label"]]
    
        
    out[index, "subject.id"] <- subject.id
    out[index, "subject.label"] <- subject.label
  }
  return(out)
}


get_object_labels <- function(path){
  
  lines <- readLines(path, ok = T, warn = F)
  
  N <- length(lines)
  
  obj_labels <- c() 

  for(index in 1:N){
    
    line <- lines[index]
    
    json.data <- fromJSON(line)
    
    obj_labels <- c(obj_labels, c(json.data[["obj_labels"]], unlist(json.data[["obj_aliases"]])))
    
  }
  return(unique(obj_labels))
  
}

get_LCP_on_all_test_data <- function(path){
  
  lines <- readLines(path, ok = T, warn = F)
  
  N <- length(lines)
  
  out <- data.frame() 

  for(index in 1:N){
    
    line <- lines[index]
    
    json.data <- fromJSON(line)
    
    subject.label <- json.data[["sub_label"]]
    
    obj_labels <- c(json.data[["obj_labels"]], unlist(json.data[["obj_aliases"]]))
    
    v_LCP <- c()
    for (obj.label in obj_labels) {
      v_LCP <- c(v_LCP, compute_pred_LCP(subject.label, obj.label))
    }
    
    out <- rbind(out, data.frame(subject.label = subject.label, obj.label = obj_labels, LCP = v_LCP))
    
  }
  
  return(out)
  
}

```

# Loading data

Please, consider that the indexing of the manual prompt (manual1 or manual2) is **inversed for P703** in the analysis compared to what is indicated in the prompt files.
For practical reasons when computing the predictings, manual1 is set to "The fungus [X] produces the compound [Y]." for rP703 and "The compound [X] was isolated and identified from culture of the fungus [Y]." for P703. Also, manual2 is set to "The compound [Y] was isolated and identified from culture of the fungus [X]." for rP703 and "The compound [X] was isolated and identified from culture of the fungus [Y]." for P703.
However, as it more convenient to group them by size of the inputed prompts during the interpretation, we reverse we associate the prompt for P703 of manual1 to manual2 and vice-versa.


## Best manual-prompting results

```{r}
manual_model_ref <- c("P703-manual*", "rP703-manual*")

# Please, set here in order the path to the results file for the BEST performing configuration with manual prompt for P703 (predicting fungi) and rP703 (predicting chemical)
manual_model_res <- c("../../output/np/manual/PubMedBERT/manual2/confidence/none/P703.json", "../../output/np/manual/PubMedBERT-full/manual1/order/confidence/rP703.json")

# Please, set here in order the path to the results file for MASK experiments for the BEST performing configuration with manual prompt for P703 (predicting fungi) and rP703 (predicting chemical)
manual_model_res_masked <- c("../../output/np/manual/PubMedBERT/manual2/confidence/none/MASKED/P703.json", "../../output/np/manual/PubMedBERT-full/manual1/order/confidence/MASKED/rP703.json")

list.of.df <- list()

for(index in 1:2){
  
  print("<=====================")
  print(manual_model_ref[index])
  p1 <- manual_model_res[index]
  p2 <- manual_model_res_masked[index]
  
  print(paste("- Std. file at: ", p1))
  print(paste("- Masked. file at: ", p2))
  
  std <- extract_prediction_proba_pairs(read_predictions(p1)) %>% distinct()
  masked <- extract_prediction_proba_pairs(read_predictions(p2)) %>% distinct()
  
  masked_for_merging <- (masked %>% select(preds.label, preds.rank))
  colnames(masked_for_merging) <- c("preds.label", "rank.in.masked")
  
  comp <- std %>% left_join(masked_for_merging, by = "preds.label")
  comp[is.na(comp$rank.in.masked), ]$rank.in.masked <- -1
  
  # fill with metadata
  comp$ref <- manual_model_ref[index]
  
  list.of.df[[index]] <- comp
  
  print("=====================>")
  
  index <- index + 1
}
          
manual_big.df <- do.call("rbind", list.of.df)

# remove empty predictions (== "") or prediction that are just a number.
manual_big.df <- manual_big.df[ (manual_big.df$preds.label != "") , ]
```



## Best OptiPrompt results

```{r}
opti_model_ref <- c("P703-opti*", "rP703-opti*")
opti_model_prop <- c("P703.json", "rP703.json")
opti_model_res_dir <- c("../../output/np/opti/PubMedBERT-full/manual1/confidence/none", "../../output/np/opti/PubMedBERT/FS/order/confidence/")
seeds <- c("seed0", "seed64", "seed333", "seed566", "seed999")
```


```{r}

j <- 1

list.of.df <- list()

for(index in 1:2){
  
  for(seed in seeds){

    print("<=====================")
    print(opti_model_ref[index])
    p1 <- paste(c(opti_model_res_dir[index], seed, opti_model_prop[index]), collapse = '/')
    p2 <- paste(c(opti_model_res_dir[index], seed, "MASKED", opti_model_prop[index]), collapse = '/')
    
    print(paste("- Std. file at: ", p1))
    print(paste("- Masked. file at: ", p2))
    
    std <- extract_prediction_proba_pairs(read_predictions(p1)) %>% distinct()
    masked <- extract_prediction_proba_pairs(read_predictions(p2)) %>% distinct()
    
    masked_for_merging <- (masked %>% select(preds.label, preds.rank))
    colnames(masked_for_merging) <- c("preds.label", "rank.in.masked")
    
    comp <- std %>% left_join(masked_for_merging, by = "preds.label")
    comp[is.na(comp$rank.in.masked), ]$rank.in.masked <- -1
    
    # comp <- compare_std_exp_to_masked(std, masked)
    
    # fill with metadata
    comp$ref <- opti_model_ref[index]
    comp$seed <- seed
    
    list.of.df[[j]] <- comp
    
    print("=====================>")
    
    j <- j + 1
  }
}
          
opti_big.df <- do.call("rbind", list.of.df)

# remove null predictions (== "") or prediction that are just a number.
opti_big.df <- opti_big.df[ (opti_big.df$preds.label != "") , ]

```

# Create plots for prompt-bias analysis

## Plots for each combination
```{r}
p1 <- create_plot(manual_big.df, "P703-manual*")
p2 <- create_plot(manual_big.df, "rP703-manual*")
p3 <- create_plot(opti_big.df, "P703-opti*")
p4 <- create_plot(opti_big.df, "rP703-opti*")
```


## Some illustrative statistics
```{r}

# For the predictions on fungus: What % of the predictions in the top 10 are common to what is predicted in the top 10 also by the model with MASK as subject, for P703:?
nrow(manual_big.df[manual_big.df$preds.rank <= 10 & manual_big.df$ref == "P703-manual*" & manual_big.df$rank.in.masked <= 10 & manual_big.df$rank.in.masked != -1 ,]) / nrow(manual_big.df[manual_big.df$preds.rank <= 10 & manual_big.df$ref == "P703-manual*",])

# For the predictions on chemicals: What % of the predictions in the top 10 are common to what is predicted by the model with MASK as subject?
nrow(manual_big.df[manual_big.df$preds.rank <= 10 & manual_big.df$ref == "rP703-manual*" & manual_big.df$rank.in.masked <= 10 & manual_big.df$rank.in.masked != -1 ,]) / nrow(manual_big.df[manual_big.df$preds.rank <= 10 & manual_big.df$ref == "rP703-manual*",])


# For the predictions on fungus: What % of the predictions in the top 10 are common to what is predicted in the top 10 also by the model with MASK as subject, for P703:?
nrow(opti_big.df[opti_big.df$preds.rank <= 10 & opti_big.df$ref == "P703-opti*" & opti_big.df$rank.in.masked <= 10 & opti_big.df$rank.in.masked != -1 ,]) / nrow(opti_big.df[opti_big.df$preds.rank <= 10 & opti_big.df$ref == "P703-opti*",])

#  For predictions on chemicals: What % of the predictions in the top 10 are common to what is predicted by the model with MASK as subject?
nrow(opti_big.df[opti_big.df$preds.rank <= 10 & opti_big.df$ref == "rP703-opti*" & opti_big.df$rank.in.masked <= 10 & opti_big.df$rank.in.masked != -1 ,]) / nrow(opti_big.df[opti_big.df$preds.rank <= 10 & opti_big.df$ref == "rP703-opti*",])
```

## Arrange all the plots
```{r}
p2 <- p2 + xlab("Prediction ranks")
p4 <- p4 + xlab("Prediction ranks")

p1 <- p1 + ylab("Proportion of predictions")
p2 <- p2 + ylab("Proportion of predictions")

full <- ggarrange(p1, p3, p2, p4, 
                  labels = c("A", "B", "C", "D"),
                  font.label = list(size = 20, color = "black", face = "bold", family = NULL), 
                  common.legend = T, 
                  legend = "right", ncol = 2, nrow = 2)
```

## Print the general plot
```{r}
plot(full)
```


# What the correct predictions are composed of ?

```{r}
test_data <- c("../../data/np/triples_processed/P703/test.jsonl", "../../data/np/triples_processed/rP703/test.jsonl")

# we get the labels of the subjets to calculate the LCP between subjects and predictions
labels_P703 <- get_table_labels(test_data[1])
labels_P703$prop <- "P703"
labels_rP703 <- get_table_labels(test_data[2])
labels_rP703$prop <- "rP703"

subject2labels <- rbind(labels_P703, labels_rP703)
```

```{r}
# We get the labels of the objects to be able to filter the predictions on labels that really correspond to chemicals or organisms
P703_obj_labels <- get_object_labels(test_data[1])
rP703_obj_labels <- get_object_labels(test_data[2])
obj_labels <- c(P703_obj_labels, rP703_obj_labels)
```

```{r}
topk <- 5
# The top answers correspond to the objects most frequently predicted by the model. We consider ONLY the predictions that really correspond to objects... not like "also" etc ...
manual_top_answers <- rbind( (manual_big.df %>% mutate(prop = str_split(ref, "-", simplify = T)[, 1]) %>% filter(prop == "P703") %>% group_by(preds.label) %>% summarise(n = n()) %>% filter(preds.label %in% tolower(obj_labels)) %>% slice_max(order_by = n, n = topk)),
                             (manual_big.df %>% mutate(prop = str_split(ref, "-", simplify = T)[, 1]) %>% filter(prop == "rP703") %>% group_by(preds.label) %>% summarise(n = n()) %>% filter(preds.label %in% tolower(obj_labels)) %>% slice_max(order_by = n, n = topk)))

opti_top_answers <- rbind( (opti_big.df %>% mutate(prop = str_split(ref, "-", simplify = T)[, 1]) %>% filter(prop == "P703") %>% group_by(preds.label) %>% summarise(n = n()) %>% filter(preds.label %in% tolower(obj_labels)) %>% slice_max(order_by = n, n = topk)),
                           (opti_big.df %>% mutate(prop = str_split(ref, "-", simplify = T)[, 1]) %>% filter(prop == "rP703") %>% group_by(preds.label) %>% summarise(n = n()) %>% filter(preds.label %in% tolower(obj_labels)) %>% slice_max(order_by = n, n = topk)))

# Frequencies are calculated for manual
manual_top_answers$avg_freq_in_preds <- NA
manual_top_answers[1:topk, ]$avg_freq_in_preds <- (manual_top_answers[1:topk, ]$n * 100) / 1934
manual_top_answers[(topk + 1):nrow(manual_top_answers), ]$avg_freq_in_preds <- (manual_top_answers[(topk + 1):nrow(manual_top_answers), ]$n * 100) / 500 

# The frequencies are calculated for opti, and here the number of seeds is taken into account:
opti_top_answers$avg_freq_in_preds <- NA
opti_top_answers[1:topk, ]$avg_freq_in_preds <- (opti_top_answers[1:topk, ]$n * 100) / (1934 * 5)
opti_top_answers[(topk + 1):nrow(manual_top_answers), ]$avg_freq_in_preds <- (opti_top_answers[(topk + 1):nrow(opti_top_answers), ]$n * 100) / (500 * 5) 


P703_freq_in_test <- get_objectlabels_freq_in_dataset(test_data[1])

# WARNING: We rename by preds.label just for the join!
colnames(P703_freq_in_test) <- c("preds.label", "n_in_test", "freq_in_test")

rP703_freq_in_test <- get_objectlabels_freq_in_dataset(test_data[2])
colnames(rP703_freq_in_test) <- c("preds.label", "n_in_test", "freq_in_test")

freq_in_test <- rbind(P703_freq_in_test, rP703_freq_in_test)
freq_in_test$preds.label <- tolower(freq_in_test$preds.label)

P703_freq_in_train <- get_objectlabels_freq_in_dataset("../../data/np/triples_processed/P703/train.jsonl")
colnames(P703_freq_in_train) <- c("preds.label", "n_in_train", "freq_in_train")

rP703_freq_in_train <- get_objectlabels_freq_in_dataset("../../data/np/triples_processed/rP703/train.jsonl")
colnames(rP703_freq_in_train) <- c("preds.label", "n_in_train", "freq_in_train")

freq_in_train <- rbind(P703_freq_in_train, rP703_freq_in_train)
freq_in_train$preds.label <- tolower(freq_in_train$preds.label)

manual_top_answers <- manual_top_answers %>% left_join(freq_in_test)
opti_top_answers <- opti_top_answers %>% left_join(freq_in_test) %>% left_join(freq_in_train)
                                                   
```

```{r}
DT::datatable(manual_top_answers)
```

```{r}
DT::datatable(opti_top_answers)
```

* Compute Longuest common prefix + generate figure
```{r}

get_class_of_preds <- function(data, subject2labels, top_answers, LCP_LIMIT){
  
  data <- data %>% mutate(prop = str_split(ref, "-", simplify = T)[, 1])
  data_for_table <- data %>% filter(is.correct) %>% left_join(subject2labels, by = c("subject.id", "prop"))
  
  
  data_for_table$LCP <- apply((data_for_table %>% select(subject.label, preds.label)), 1, FUN = function(x){compute_pred_LCP(x[1], x[2])})
  data_for_table$test_LCP <- as.integer(data_for_table$LCP >= LCP_LIMIT)
  
  data_for_table$test_freq <- as.integer(data_for_table$preds.label %in% top_answers$preds.label)
  
  data_for_table <- data_for_table %>% mutate(Class = paste0(test_freq, test_LCP))
  data_for_table$Class <- factor(data_for_table$Class, levels = c("00", "10", "01", "11"), labels = c("Original", "Top-Freq", "Leak", "Top-Freq & Leak"))
  
  return(list(data_for_table, (data_for_table %>% group_by(prop, Class) %>% summarise(n = n()) %>% group_by(prop) %>% mutate(p = round(n / sum(n) * 100, 2) ))))

}

LCP_LIMIT <- 5
manual <- get_class_of_preds(manual_big.df, subject2labels, manual_top_answers, LCP_LIMIT)
opti <- get_class_of_preds(opti_big.df, subject2labels, opti_top_answers, LCP_LIMIT)

data_for_plot <- rbind( (manual[[2]] %>% mutate(Exp = "Manual*")), (opti[[2]] %>% mutate(Exp = "Opti*")) )
data_for_plot <- data_for_plot %>% mutate(label = paste("(", p, " %)"))

data_for_plot %>% 
  ggplot(aes(x = Exp, y = p, fill = Class, label = label)) + 
  geom_bar(stat = "identity", position = "stack") +
  theme_classic() +
  geom_text(size = 6, position = position_stack(vjust = 0.5)) +
  theme(axis.text = element_text(size = 20, colour="black"), text = element_text(size = 15, colour="black")) +
  facet_grid(. ~ prop) +
    scale_fill_brewer(palette="Set1")

data_for_plot %>% group_by(prop, Exp) %>% summarise(N = sum(n))
```

Some examples of leaks: (we only do this for Opti because it doesn't change much)

```{r}
opti[[1]] %>% filter(Class == "Leak", prop == "P703") %>% sample_n(10) %>% DT::datatable()
```

```{r}
opti[[1]] %>% filter(Class == "Leak", prop == "rP703") %>% sample_n(10) %>% DT::datatable()
```

We calculate the global stats of the LCP

```{r}
P703_test_LCP <- get_LCP_on_all_test_data(test_data[1])

print(paste("Considering all the object labels and alias, the proportion of LCP < 5 in the test set for P703 is: ", sum(P703_test_LCP$LCP < 5) / nrow(P703_test_LCP) * 100))

rP703_test_LCP <- get_LCP_on_all_test_data(test_data[2])
print(paste("Considering all the object labels and alias, the proportion of LCP < 5 in the test set for P703 is: ", sum(rP703_test_LCP$LCP < 5) / nrow(rP703_test_LCP) * 100))
```

